

\chapter{Calibration Glossary}
\shwlabel{ChapterGlossary}

\begin{description}

\item[Axis]
  Each rope or umbilical in the manipulator system is referred to
  as an axis.

\item[Data Concentrator]
  Is the box that reads out the encoder boards that measure the 
  manipulator rope lengths and tensions.


\item[DCR]
  Deck Clean Room.

\item[Encoder Box]
  Each manipulator {\em Axis} has an encoder box to which is attached
  a shaft encoder to measure the axis length and a load cell to measure
  the tension of the axis.  The shaft encoder and load cell are
  readout through the encoder box by the {\em Data Concentrator}.

\item[Glovebox]
  The box mounted on the Universal Interface over top of the 
neck of the detector.  Calibration sources that are deployed into
the detector are mounted in URMs which are attached to the glovebox
at either gate valve 1 (10 inch) or gate valve 3 (4 inch).  There
is also a 6 inch gate valve which is not used.

\item[Manipulator] The systems used to deploy calibration sources
  into the SNO detector.

\item[Polyaxis] is the name used in the manipulator system
  for multiple axes (ropes or umbilicals) attached to a calibration
  source.

\item[Umbilical]
  The umbilical is a 1/2 inch diameter silicone tube that which
  carries the services to the source.

\item[URM]
  Umbilical Retreival Mechanism.  The sources are mounted in
  {\em Source Tubes} on URMs.  The URM contains a rope which
  supports the weight of the source and an {\em Umbilical} which
  carries the services to the source.

\item[Watchdog Timer Box]
  The manipulator stepper motors are controlled by a National Instruments
  TIO10 card.  The Watchdog Timer Box is actually a breakout box for
  the TIO10 card.  The manipulator computer periodically sends a 
  {\em motor enable} signal to the box.  If this signal is not present
  the manipulator motors are shut down.



\end{description}
