\documentclass{article}
\begin{document}

\center{\emph{Sudbury Neutrino Observatory}} \\   
\center{\textbf{Procedure for Changing from Laserball to Acrylic Source.}} \\
\vspace{7mm}
\begin{tabular}{lc}                                                
\textbf{Authors:}  Fraser Duncan, Will Peters    &    \textbf{Date Created:}  3 Dec. 2001   \\ 
                              &                                \\ 
\textbf{Operator(s):}       &           \textbf{Date:} \\
\end{tabular}

   \begin{enumerate}
      
\checkitem Remove salt probe / plug clamp. 
\checkitem While supporting both laserball and can remove screws which hold the laserball to the can.
\checkitem Gently slide the laserball away from the can until CAJON fitting is exposed. Be sure to support both
can and laserball and make sure they're aligned. 
\checkitem Carefully unscrew top end of CAJON fitting and pull out the fiber end. DO not let it retract into the can.
\checkitem Store laserball in toolbox
\checkitem Ensure o-ring is in place on stem.
\checkitem Ensure dummy CAJON fitting is pin place on stem
\checkitem Carefully insert fiber end into stem CAJON fitting and tighten lightly.
\checkitem Verify the saltprobe / plug o-ring is still in place.
\checkitem Push the fiber in to the cajon fitting on the stem. Tighten the fitting ligthly.
\checkitem Push the stem up against the can. Make sure the fiber is not bent or pinched
in any way.
\checkitem Slide the steel ring into place and bolt plate plus stem to can. Use correct screws.
\checkitem Slide clamp up around saltprobe or plug. Make sure it is oriented correctly ( so that
 it fits ) and
tighten screw appropriately. ( {\bf Do not overtigthen  } ! )

  The next few steps depend on the desired source configuration ( canned, un-canned, 
un-canned with spacer, blind can ).


\checkitem Canned source
\begin{enumerate}


\end{enumerate}

\checkitem Uncanned source
\begin{enumerate}


\end{enumerate}

\checkitem Uncanned source, spacer
\begin{enumerate}


\end{enumerate}


   \end{enumerate}


\end{document}
